### 2017-02-27 Theory/Theory.tex
More editing

The quantum number associated with the strong force is colour, which
both the quarks and the gluons carry.
Within the framework of \ac{QFT}, the
strong force is described by \ac{QCD}, governed by an SU(3) (colour) symmetry.
%There are 8 gluons due to the SU(3) symmetry, colour octet and a colour singlet, but
%if there was a colour singlet it should occur as a free particle while it does not.
%See chapter 285 of griffiths
Characteristics of \ac{QCD} include confinement, meaning quarks are always observed in bound
states called hadrons, and asymptotic
freedom \cite{asympt-I,asympt-II}. Due to asymptotic freedom the 
strong coupling strength decreases at 
high energy (or short distances), making the force weaker. 
A more detailed introduction to \ac{QCD} can 
be found in reference \cite{griffiths}.

@@

The associated gauge fields are the three weak isospin fields $W_{\mu}^{i}$ and the
weak hypercharge field $B_{\mu}$. These four fields mix to form the physical \Pphoton,
\PZ and $\PW^{\pm}$ states as
\begin{equation}\label{eqn:ewk_propagators_W}
\PW^{\pm} &= \frac{1}{\sqrt{2}}(W_{\mu}^1 \mp i W_{\mu}^2),
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_phot}
A_{\mu} &= W_{\mu}^3\sin{\theta_W} + B_{\mu}\cos{\theta_W},
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_Z}
Z_{\mu} &= W_{\mu}^3\cos{\theta_W} - B_{\mu}\sin{\theta_W}.
\end{equation}
In these equations $A_{\mu}$ is the photon field and $Z_{\mu}$ the \PZ field.
The weak mixing angle $\theta_W$ is related to the weak (g) and electromagnetic (g')
coupling constants as
\begin{equation}\label{eqn:thetaw}
\begin{split}
&g\sin{\theta_W} = g'\cos{\theta_W}\\
&\Rightarrow \sin{\theta_W} = \frac{g'}{\sqrt{g^2+g'^2}} \text{ and } \cos{\theta_W} = \frac{g}{\sqrt{g^2+g'^2}}
\end{split}
\end{equation}

@@

\subsection{The Higgs mechanism}
\label{sec:theory_sm_higgsmech}
The mechanism of electroweak symmetry breaking, the
Higgs mechanism, was proposed in the 1960s by Englert and Brout \cite{englertbrout},
Higgs \cite{higgs-I,higgs-II,higgs-III} and Guralnik, Hagen and Kibble \cite{GHK}.
The idea is to add a field that is symmetric under gauge transformations, but
has a nonzero vacuum expectation value and thus breaks the symmetry spontaneously.
According to Goldstone's theorem \cite{goldstone-theorem,goldstone-theorem-2} a 
consequence of spontaneous symmetry breaking is the appearance of a 
massless scalar (Goldstone boson) for each broken generator.

@@

Under a local $SU(2)\times U(1)$ transformation, the field transforms as,
\begin{equation}\label{eqn:hfield_trsf}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix} = \exp{\{\frac{i}{2}(\vec{\theta}\cdot\vec{\sigma} + \rho)\}}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix}, 
\end{equation}
where $\vec{\theta}$ and $\rho$
are phase transformations which are functions of the space-time coordinates. The covariant
derivative associated with such a transformation is
\begin{equation}\label{eqn:H_cov_deriv}
D_{\mu} = \partial_{\mu} - \frac{i}{2}g\vec{W_{\mu}}\cdot\vec{\sigma} - \frac{i}{2}g'B_{\mu}.
\end{equation}
In the vacuum state, $V(\Phi^{\dagger}\Phi) =0$, $\Phi^{\dagger}\Phi=\frac{\mu_{\text{SM}}^2}{2\lambda}$ and so
$\Phi$ has to be nonzero.
Now choosing $\phi^+$ = 0 and taking $\phi^0$ to be real, this means:
%\begin{equation}\label{eqn:vacuum_expt}
%\begin{split}
%&\Phi^{\dagger}\Phi - \frac{\mu^2}{2\lambda} = 0 \Rightarrow \\
%&\Phi^{\dagger}\Phi = \frac{\mu^2}{2\lambda} \Rightarrow\\
%&\phi^{+*}\phi^+ + \phi^{0*}\phi^0 = \frac{\mu_^2}{2\lambda} \text{ and choosing } \phi^+ \text{ = 0 and } \phi^0 \text{ to be real,}\\
%&\phi^0 = \frac{v}{\sqrt{2}} \text{ where } v \text{ is real and} v^2=\frac{\mu^2}{\lambda}.
%\end{split}
%\end{equation}
\begin{equation}\label{eqn:field_vev}
\langle 0 | \Phi | 0 \rangle = \begin{pmatrix} 0 \\
\frac{v}{\sqrt{2}} \end{pmatrix},
\end{equation}
where $v$ is a real parameter, $v^2 \equiv \frac{\mu_{\text{SM}}^2}{\lambda}$.

Considering infinitesimal fluctuations around the vacuum state, it can be shown that the
only remaining generator that leaves the field invariant is $I_3+\frac{Y}{2}$, the generator
of the U(1) group.
%\begin{equation}\label{eqn:fluctvac}
%\begin{split}
%&\delta\Phi = \frac{i}{2}(\vec{\theta}\cdot\vec{\sigma} + \rho)\Phi\\
%&=\frac{i}{2}\begin{pmatrix} \theta_3 + \rho & \theta_1 - i\theta_2 \\
%\theta_1+i\theta_2 & \rho - \theta_3 \end{pmatrix} \begin{pmatrix} 0\\
%\frac{v}{\sqrt{2}} \end{pmatrix}.
%\end{split}
%\end{equation}
%Requiring $\delta\Phi=0$, we have $\theta_1=\theta_2=0$ and $\rho=\theta_3$, 
%and the only generator that leaves the field invariant is $I_3+Y$, the generator of the U(1) group.
Denoting the 
remaining, broken, generators as $\vec{b}$, and parameterising the field as an expansion around the
vacuum state gives
\begin{equation}\label{eqn:vev_expansion}
\Phi = \exp{\{\frac{i}{\sqrt{2}v}\vec{\theta}\cdot\vec{b}\}}\begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}} (v+\PHiggs) \end{pmatrix}.
\end{equation}
By choosing an appropriate gauge the Goldstone bosons $\vec{\theta}$
can be eliminated. They become the longitudinal degrees of freedom of the $\PW^{\pm}$ and \PZ 
%longitudinal polarisation: ms=0
bosons. This then leaves
\begin{equation}\label{eqn:vev_expansion_satisfied}
\Phi = \begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}}(v+\PHiggs) \end{pmatrix}.
\end{equation}
Substituting this into the Lagrangian of equation \ref{eqn:higgs_gen_lag}, 
and using the definitions of $\PW_{\mu}^{\pm}$ and $\PZ_{\mu}$ from 
equations \ref{eqn:ewk_propagators_W}-\ref{eqn:ewk_propagators_Z}, the Lagrangian becomes
\begin{equation}\label{eqn:completed_lagrangian}
\begin{split}
&\mathcal{L}_{\text{Higgs}} = -\frac{1}{2}\partial_{\mu}\PHiggs\partial^{\mu}\PHiggs -\frac{g^2v^2}{4}\PW_{\mu}^- \PW^{\mu +} - \frac{g^2+g'^2}{8}v^2\PZ_{\mu}\PZ^{\mu}  - \mu_{\text{SM}}^2\PHiggs^2  + \\ 
&\text{terms of at least }\mathcal{O}(3)\text{ in the fields}.
\end{split}
\end{equation}
This Lagrangian contains mass terms for the $\PW^{\pm}$ and \PZ bosons, with 
$m_{\PW} = \frac{gv}{2}$ and $m_{\PZ} = \frac{1}{2}v\sqrt{g^2+g'^2}$. Using equation \ref{eqn:thetaw} 
we can see that these two masses are
related as $\frac{m_{\PW}}{m_{\PZ}} = \cos{\theta_W}$. The remaining mass term is for the Higgs field \PHiggs, 
$M_{\PHiggs} = \sqrt{2\mu_{\text{SM}}^2}$. It is denoted $M_{\PHiggs}$ to avoid confusion with the mass of the heavy scalar
\PHiggs boson in the \ac{MSSM} $m_{\PHiggs}$, which will be introduced in section \ref{sec:theory_MSSM_H}.
Note that the photon does not acquire a mass.
%Mass term : -1/2m_X^2. However for m_W note W_mu+Wmu- = 1/2(W_mu^1+iW_mu^2)(W_mu^1-iW_mu^2) = 
%1/2*(W_mu^1W_mu^1+W_mu^2W_mu^2). But also W+^2+W^-2 = W1^2+W2^2 so masses end up the same

Using these developments it is also possible to add mass terms for the fermions
through Yukawa couplings between the fermion- and Higgs fields, which are of the form
\begin{equation}\label{eqn:yukawa_coupl}
\lambda_f(\bar{\psi_L}\Phi\psi_R + \bar{\psi_R}\Phi\psi_L).
\end{equation}
The parameter $\lambda_f$ is a coupling constant specific to each fermion, $\lambda_f \propto m_f$.
This means that heavier fermions couple more strongly to the Higgs field.

\section{Standard model Higgs boson measurements}
\label{sec:theory_smH}
The mass of the Higgs boson is a free parameter in the \ac{SM}, and so
a large range of possible masses needed to be covered in Higgs boson searches. 
Searches for the Higgs boson at the \ac{LEP} did not lead
to its observation. However, masses below $114.4\,\GeV$ were excluded at 95\% \ac{CL} \cite{LEP-Higgs},
providing a lower search limit. Subsequent searches at the 
Tevatron also did not lead to discovery of the Higgs boson, but did exclude 
an additional mass range of $140$--$186\,\GeV$ at 95\% \ac{CL} \cite{TEV-Higgs}.

On the fourth of July 2012 the ATLAS and CMS Collaborations
announced the discovery of a boson with a mass of around $125\,\GeV$ \cite{HDiscoveryATLAS,HDiscoveryCMS}.
The discovery was based on the observation of excesses in the 
search for $\PHiggs\rightarrow\gamma\gamma$ and $\PHiggs\rightarrow \PZ\PZ\rightarrow 4\ell$, with a significance 
above 5$\sigma$.
More data were collected and analysed during the remainder of 2012, and subsequent studies of its properties,
such as spin, parity \cite{ATLASspin,CMSspin} and couplings to other particles, 
increased the confidence in the compatibility of this new state with the \ac{SM} Higgs boson. 
The most precise measurement of the mass of the Higgs boson, $M_{\PHiggs} = 125.09 \pm 0.21 \text{ (stat) } \pm 0.11 \text{ (syst)}\,\GeV$, 
comes from the ATLAS and CMS combined
mass measurement using the full Run 1 dataset \cite{MassComb}.
The combined ATLAS and CMS production- and decay rate measurements, also using the full Run 1 dataset,
show very good agreement with the \ac{SM} predictions \cite{CouplComb}. In addition, 
the $\PHiggslight \rightarrow \Pgt\Pgt$ decay was observed with a significance of 5.5$\sigma$ through this combination. During Run 1
searches by the ATLAS and CMS Collaborations resulted in observed 
significances of $4.5\sigma$ \cite{ATLAS-tautau} and $3.2\sigma$ \cite{SMHtautauCMS}, respectively. 
%Despite the good agreement with the \ac{SM}, some discrepancies do remain, 
%and so the measurements of \ac{SM} Higgs boson

@@

Figure \ref{fig:theory_smhprod} shows Feynman diagrams of the dominant Higgs boson production modes
at the \ac{LHC}. From the production cross sections for each mode, shown in figure
\ref{fig:theory_smhxsbr}a, we see that gluon fusion production is by far dominant. The other modes
shown have cross sections at least an order of magnitude smaller, but are of importance due to their topologies.
Tagging the additional leptons or jets in the final states of
these production modes can reduce the size
of the \ac{SM} background.

@@

Considering the branching ratios of the \ac{SM} Higgs boson in figure \ref{fig:theory_smhxsbr}b, we observe that
the branching ratios of
$\PHiggs \rightarrow \gamma\gamma$ and $\PHiggs \rightarrow \PZ\PZ$, the two discovery
channels, are smaller than the branching ratios into many other final states.
These two channels are thus not sensitive as a result of the size of their branching ratio, but
rather due to the small backgrounds in these analyses as well as the excellent photon and lepton energy resolution the detectors provide.

During Run 2 of the LHC, \ac{SM} Higgs boson measurements, and searches for the 
\ac{SM} Higgs boson in decay channels not previously established, have 
already been performed. The decay in to $\gamma\gamma$ has been re-established \cite{CMSHgamgam2016,ATLASHgamgam2016},
as has the decay in to $\PZ\PZ\rightarrow 4\ell$ \cite{CMSHZZ2016,ATLASHZZ2016}, and measurements
of $\PHiggs\rightarrow WW$ are being performed \cite{CMSHWW2016,ATLASHWW2016}. Establishing evidence for the decay of 
$\PHiggs\rightarrow \Pbottom\Pbottom$ \cite{CMSVBFHbb2016,ATLASVHbb2016} and of $\PHiggs\rightarrow \mu\mu$ \cite{ATLASHmm2016} 
is still to be achieved.
Finally, searches for $\Ptop\Ptop\PHiggs$ production, which are important to probe directly the top-Higgs coupling, have also been carried 
out \cite{CMSttH2016,CMSttHmultilep2016,ATLASttHbb2016,ATLASttHmultilep2016}. As the Run 2 dataset
keeps growing studies at the \acs{LHC} will continue to test the compatibility of the Higgs boson with the \ac{SM}.

\section{Beyond the standard model}
\label{sec:theory_BSM}
The \ac{SM} has been successfully tested to high accuraccy \cite{pdg-2014}. However,
some issues remain that cannot be addressed by the \ac{SM} alone \cite{griffiths}.
For example, it does not provide a candidate for dark matter, which is estimated
to make up 25\% of the matter energy density in the universe.
In addition, the running coupling constants of the electroweak and
strong forces do not intersect at a common energy scale. Another problem is that the force
of gravity does not appear in the \ac{SM}.

These points aside, the most important issue concerning
the Higgs sector of the \ac{SM} is the hierarchy problem.
The \ac{SM}
is accepted to be an effective field theory that describes
the elementary particles and their interactions well, up to an energy scale $\Lambda$ beyond which it breaks down.
This is the energy scale at which new physics must enter.
We know that $\Lambda \leq \mathcal{O}(10^{19})$, the Planck scale, as quantum 
gravitational effects start to become important in that region. The energy scale $\Lambda$ enters
the corrections due to fermion and boson loops to the Higgs boson mass quadratically \cite{MSSM-carena-haber}:
\begin{equation}\label{eqn:mh_hierarchy}
\begin{split}
&M_{H_{\text{SM}}}^2  = (M_H^2)_0 + \Delta M_H^2,\\
&\Delta M_H^2 \sim \mathcal{O}(\Lambda^2).
\end{split}
\end{equation}
Thus assuming that there is no new physics all the way up to
the Planck scale the observation $M_{\PHiggs}=125\,\GeV$ would only be compatible with $\Lambda$ of order $10^{19}\,\GeV$
if there was extreme fine-tuning of the bare Higgs boson mass $(M_{\PHiggs}^2)_0$.

@@

Using this we can define

@@

The remaining 5 degrees of freedom lead to five physical Higgs bosons, a charged

@@

and the neutral scalars \PHiggslight and \PHiggs. These arise from mixing of the

@@

m_{\PHiggsps}^2\sin{\beta}^2 + m_{\PZ}^2\cos{\beta}^2 & -(m_{\PHiggsps}^2+m_{\PZ}^2)\sin{\beta}\cos{\beta}\\
-(m_{\PHiggsps}^2+m_{\PZ}^2)\sin{\beta}\cos{\beta} & m_{\PHiggsps}^2\cos{\beta}^2+m_{\PZ}^2\sin{\beta}^2 \end{pmatrix},

@@

m_{\PHiggs,\PHiggslight}^2 = \frac{1}{2}[m_{\PHiggsps}^2+m_{\PZ}^2 \pm \sqrt{(m_{\PHiggsps}^2+m_{\PZ}^2)^2-4m_{\PZ}^2m_{\PHiggsps}^2\cos{2\beta}^2}].

@@

\cos{(\beta-\alpha)}^2 = \frac{m_{\PHiggslight}^2(m_{\PZ}^2-m_{\PHiggslight}^2)}{m_{\PHiggsps}^2(m_{\PHiggs}^2-m_{\PHiggslight}^2)}.

@@

The tree-level couplings of the neutral Higgs bosons to bosons and fermions 
are modified with respect to the couplings of the Higgs boson in the \ac{SM}, by the multiplicative factors
given in table \ref{tab:mssm_couplings} \cite{YR4}.

@@

The dominant neutral MSSM Higgs boson production processes at hadron colliders
are slightly different from the \ac{SM} Higgs boson production processes. 
Because the pseudoscalar A does not couple to the
vector bosons, and in the decoupling limit the coupling of \PHiggs to vector
bosons is suppressed with respect to the \ac{SM} expectation, VBF and \PW or \PZ associated
production are not as important as for the production of the \ac{SM} Higgs boson. As in the \ac{SM}, gluon fusion production is 
the dominant production mode at low \tanb.
An example tree-level Feynman diagram for gluon fusion production, which proceeds via
a quark loop, is given in figure \ref{fig:production_mssm}a.
At low \tanb~values this loop is dominated by top quarks, while at 
high \tanb~the b-quark loop dominates. This is a result of the enhanced coupling to down-type fermions at high \tanb. 
Due to the negative top-bottom loop interference effect the gluon fusion cross section decreases with increasing \tanb~up
to around \tanb~$=3$, when the cross section starts to increase with increasing \tanb~again.

>>>>

The quantum number associated with the strong force is colour, which
both the quarks and the gluons carry.
Within the framework of \ac{QFT}, the
strong force is described by \ac{QCD}, governed by an SU(3) (colour) symmetry.
%There are 8 gluons due to the SU(3) symmetry, colour octet and a colour singlet, but
%if there was a colour singlet it should occur as a free particle while it does not.
%See chapter 285 of griffiths
Characteristics of \ac{QCD} include confinement, meaning quarks are always observed in bound
states called hadrons, and asymptotic
freedom \cite{asympt-I,asympt-II}. Due to asymptotic freedom the 
strong coupling strength decreases at 
high energy, or short distances, making the force weaker. 
A more detailed introduction to \ac{QCD} can 
be found in reference \cite{griffiths}.

@@

The associated gauge fields are the three weak isospin fields $W_{\mu}^{i}$ and the
weak hypercharge field $B_{\mu}$. These four fields mix to form the physical \Pphoton,
\PZ and $\PW^{\pm}$ states as
\begin{equation}\label{eqn:ewk_propagators_W}
\PW_{\mu}^{\pm} &= \frac{1}{\sqrt{2}}(\PW_{\mu}^1 \mp i \PW_{\mu}^2),
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_phot}
\text{A}_{\mu} &= \PW_{\mu}^3\sin{\theta_W} + B_{\mu}\cos{\theta_W},
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_Z}
\PZ_{\mu} &= \PW_{\mu}^3\cos{\theta_W} - \text{B}_{\mu}\sin{\theta_W}.
\end{equation}
In these equations $\text{A}_{\mu}$ is the photon field and $\PZ_{\mu}$ the \PZ field.
The weak mixing angle $\theta_W$ is related to the weak (g) and electromagnetic (g')
coupling constants as
\begin{equation}\label{eqn:thetaw}
\begin{split}
&g\sin{\theta_W} = g'\cos{\theta_W}\\
&\Rightarrow \sin{\theta_W} = \frac{g'}{\sqrt{g^2+g'^2}} \text{ and } \cos{\theta_W} = \frac{g}{\sqrt{g^2+g'^2}}
\end{split}
\end{equation}

@@

\subsection{The Higgs mechanism}
\label{sec:theory_sm_higgsmech}
The mechanism of electroweak symmetry breaking, the
Higgs mechanism, was proposed in the 1960s by Englert and Brout \cite{englertbrout},
Higgs \cite{higgs-I,higgs-II,higgs-III} and Guralnik, Hagen and Kibble \cite{GHK}.
The idea is to add a field that is symmetric under gauge transformations, but
has a non-zero vacuum expectation value and thus breaks the symmetry spontaneously.
According to Goldstone's theorem \cite{goldstone-theorem,goldstone-theorem-2} a 
consequence of spontaneous symmetry breaking is the appearance of a 
massless scalar (Goldstone boson) for each broken generator.

@@

Under a local SU(2)$\times$U(1) transformation, the field transforms as,
\begin{equation}\label{eqn:hfield_trsf}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix} = \exp{\{\frac{i}{2}(\vec{\theta}\cdot\vec{\sigma} + \rho)\}}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix}, 
\end{equation}
where $\vec{\theta}$ and $\rho$
are phase transformations which are functions of the space-time coordinates. The covariant
derivative associated with such a transformation is
\begin{equation}\label{eqn:H_cov_deriv}
D_{\mu} = \partial_{\mu} - \frac{i}{2}g\vec{\PW_{\mu}}\cdot\vec{\sigma} - \frac{i}{2}g'\text{B}_{\mu}.
\end{equation}
In the vacuum state, $V(\Phi^{\dagger}\Phi) =0$, $\Phi^{\dagger}\Phi=\frac{\mu_{\text{SM}}^2}{2\lambda}$ and so
$\Phi$ has to be non-zero.
Now choosing $\phi^+$ = 0 and taking $\phi^0$ to be real, this means:
%\begin{equation}\label{eqn:vacuum_expt}
%\begin{split}
%&\Phi^{\dagger}\Phi - \frac{\mu^2}{2\lambda} = 0 \Rightarrow \\
%&\Phi^{\dagger}\Phi = \frac{\mu^2}{2\lambda} \Rightarrow\\
%&\phi^{+*}\phi^+ + \phi^{0*}\phi^0 = \frac{\mu_^2}{2\lambda} \text{ and choosing } \phi^+ \text{ = 0 and } \phi^0 \text{ to be real,}\\
%&\phi^0 = \frac{v}{\sqrt{2}} \text{ where } v \text{ is real and} v^2=\frac{\mu^2}{\lambda}.
%\end{split}
%\end{equation}
\begin{equation}\label{eqn:field_vev}
\langle 0 | \Phi | 0 \rangle = \begin{pmatrix} 0 \\
\frac{v}{\sqrt{2}} \end{pmatrix},
\end{equation}
where $v$ is a real parameter, $v^2 \equiv \frac{\mu_{\text{SM}}^2}{\lambda}$.

Considering infinitesimal fluctuations around the vacuum state, it can be shown that the
only remaining generator that leaves the field invariant is $I_3+\frac{Y}{2}$, the generator
of the U(1) group.
%\begin{equation}\label{eqn:fluctvac}
%\begin{split}
%&\delta\Phi = \frac{i}{2}(\vec{\theta}\cdot\vec{\sigma} + \rho)\Phi\\
%&=\frac{i}{2}\begin{pmatrix} \theta_3 + \rho & \theta_1 - i\theta_2 \\
%\theta_1+i\theta_2 & \rho - \theta_3 \end{pmatrix} \begin{pmatrix} 0\\
%\frac{v}{\sqrt{2}} \end{pmatrix}.
%\end{split}
%\end{equation}
%Requiring $\delta\Phi=0$, we have $\theta_1=\theta_2=0$ and $\rho=\theta_3$, 
%and the only generator that leaves the field invariant is $I_3+Y$, the generator of the U(1) group.
Denoting the 
remaining, broken, generators as $\vec{b}$, and parameterising the field as an expansion around the
vacuum state gives
\begin{equation}\label{eqn:vev_expansion}
\Phi = \exp{\{\frac{i}{\sqrt{2}v}\vec{\theta}\cdot\vec{b}\}}\begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}} (v+\PHiggs) \end{pmatrix}.
\end{equation}
By choosing an appropriate gauge the Goldstone bosons $\vec{\theta}$
can be eliminated. They become the longitudinal degrees of freedom of the $\PW^{\pm}$ and \PZ 
%longitudinal polarisation: ms=0
bosons. This then leaves
\begin{equation}\label{eqn:vev_expansion_satisfied}
\Phi = \begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}}(v+\PHiggs) \end{pmatrix}.
\end{equation}
Substituting this into the Lagrangian of equation \ref{eqn:higgs_gen_lag}, 
and using the definitions of $\PW_{\mu}^{\pm}$ and $\PZ_{\mu}$ from 
equations \ref{eqn:ewk_propagators_W}--\ref{eqn:ewk_propagators_Z}, the Lagrangian becomes
\begin{equation}\label{eqn:completed_lagrangian}
\begin{split}
&\mathcal{L}_{\text{Higgs}} = -\frac{1}{2}\partial_{\mu}\PHiggs\partial^{\mu}\PHiggs -\frac{g^2v^2}{4}\PW_{\mu}^- \PW^{\mu +} - \frac{g^2+g'^2}{8}v^2\PZ_{\mu}\PZ^{\mu}  - \mu_{\text{SM}}^2\PHiggs^2  + \\ 
&\text{terms of at least }\mathcal{O}(3)\text{ in the fields}.
\end{split}
\end{equation}
This Lagrangian contains mass terms for the $\PW^{\pm}$ and \PZ bosons, with 
$m_{\PW} = \frac{gv}{2}$ and $m_{\PZ} = \frac{1}{2}v\sqrt{g^2+g'^2}$. Using equation \ref{eqn:thetaw} 
we can see that these two masses are
related as $\frac{m_{\PW}}{m_{\PZ}} = \cos{\theta_W}$. The remaining mass term is for the Higgs field \PHiggs, 
$M_{\PHiggs} = \sqrt{2\mu_{\text{SM}}^2}$. It is denoted $M_{\PHiggs}$ to avoid confusion with the mass of the heavy scalar
\PHiggs boson in the \ac{MSSM} $m_{\PHiggs}$, which will be introduced in section \ref{sec:theory_MSSM_H}.
Note that the photon does not acquire a mass.
%Mass term : -1/2m_X^2. However for m_W note W_mu+Wmu- = 1/2(W_mu^1+iW_mu^2)(W_mu^1-iW_mu^2) = 
%1/2*(W_mu^1W_mu^1+W_mu^2W_mu^2). But also W+^2+W^-2 = W1^2+W2^2 so masses end up the same

Using these developments it is also possible to add mass terms for the fermions
through Yukawa couplings between the fermion and Higgs fields, which are of the form
\begin{equation}\label{eqn:yukawa_coupl}
\lambda_f(\bar{\psi_L}\Phi\psi_R + \bar{\psi_R}\Phi\psi_L).
\end{equation}
The parameter $\lambda_f$ is a coupling constant specific to each fermion, $\lambda_f \propto m_f$.
This means that heavier fermions couple more strongly to the Higgs field.

\section{Standard model Higgs boson measurements}
\label{sec:theory_smH}
The mass of the Higgs boson is a free parameter in the \ac{SM}, and so
a large range of possible masses needed to be covered in Higgs boson searches. 
Searches for the Higgs boson at the \ac{LEP} collider did not lead
to its observation. However, masses below $114.4\,\GeV$ were excluded at the 95\% \ac{CL} \cite{LEP-Higgs},
providing a lower search limit. Subsequent searches at the 
Tevatron also did not lead to discovery of the Higgs boson, but did exclude 
an additional mass range of $140$--$186\,\GeV$ at 95\% \ac{CL} \cite{TEV-Higgs}.

On the fourth of July 2012 the ATLAS and CMS Collaborations
announced the discovery of a boson with a mass of around $125\,\GeV$ \cite{HDiscoveryATLAS,HDiscoveryCMS}.
The discovery was based on the observation of excesses in the 
search for $\PHiggs\rightarrow\gamma\gamma$ and $\PHiggs\rightarrow \PZ\PZ\rightarrow 4\ell$, with a significance 
above 5$\sigma$.
More data were collected and analysed during the remainder of 2012, and subsequent studies of its properties,
such as spin, parity \cite{ATLASspin,CMSspin} and couplings to other particles, 
increased the confidence in the compatibility of this new state with the \ac{SM} Higgs boson. 
The most precise measurement of the mass of the Higgs boson, $M_{\PHiggs} = 125.09 \pm 0.21 \text{ (stat) } \pm 0.11 \text{ (syst)}\,\GeV$, 
comes from the ATLAS and CMS combined
mass measurement using the full dataset collected up to the end of 2012 \cite{MassComb}.
The combined ATLAS and CMS production and decay rate measurements, also using the full dataset collected up to the end of 2012,
show very good agreement with the \ac{SM} predictions \cite{CouplComb}. In addition, 
the $\PHiggs \rightarrow \Pgt\Pgt$ decay was observed with a significance of 5.5$\sigma$ through this combination. During the 2010--2012
LHC running period searches by the ATLAS and CMS Collaborations resulted in observed 
significances of $4.5\sigma$ \cite{ATLAS-tautau} and $3.2\sigma$ \cite{SMHtautauCMS} for this process, respectively. 
%Despite the good agreement with the \ac{SM}, some discrepancies do remain, 
%and so the measurements of \ac{SM} Higgs boson

@@

Figure \ref{fig:theory_smhprod} shows Feynman diagrams of the dominant Higgs boson production modes
at the \ac{LHC}. The production cross sections for each mode, shown in figure
\ref{fig:theory_smhxsbr}a, show that gluon fusion production is by far dominant. The other modes
shown have cross sections at least an order of magnitude smaller, but are of importance due to their topologies.
Tagging the additional leptons or jets in the final states of
these production modes can reduce the size
of the \ac{SM} background.

@@

The branching ratios of the \ac{SM} Higgs boson in figure \ref{fig:theory_smhxsbr}b show that
the branching ratios of
$\PHiggs \rightarrow \gamma\gamma$ and $\PHiggs \rightarrow \PZ\PZ$, the two discovery
channels, are smaller than the branching ratios into many other final states.
These two channels are thus not sensitive as a result of the size of their branching ratio, but
rather due to the small backgrounds in these analyses as well as the excellent photon and lepton energy resolution the detectors provide.

Since the re-start of the LHC in 2015, \ac{SM} Higgs boson measurements, and searches for the 
\ac{SM} Higgs boson in decay channels not previously established, have 
already been performed. The decay in to $\gamma\gamma$ has been re-established \cite{CMSHgamgam2016,ATLASHgamgam2016},
as has the decay in to $\PZ\PZ\rightarrow 4\ell$ \cite{CMSHZZ2016,ATLASHZZ2016}, and measurements
of $\PHiggs\rightarrow \PW\PW$ have been performed \cite{CMSHWW2016,ATLASHWW2016}. Establishing evidence for the decay of 
$\PHiggs\rightarrow \Pbottom\Pbottom$ \cite{CMSVBFHbb2016,ATLASVHbb2016} and of $\PHiggs\rightarrow \mu\mu$ \cite{ATLASHmm2016} 
is still to be achieved.
Finally, searches for $\Ptop\Ptop\PHiggs$ production, which are important to directly probe the top-Higgs coupling, have also been carried 
out \cite{CMSttH2016,CMSttHmultilep2016,ATLASttHbb2016,ATLASttHmultilep2016}. As more data
are collected studies at the \acs{LHC} will continue to test the compatibility of the Higgs boson with the \ac{SM}.

\section{Beyond the standard model}
\label{sec:theory_BSM}
The \ac{SM} has been successfully tested to high accuracy \cite{pdg-2014}. However,
some issues remain that cannot be addressed by the \ac{SM} alone \cite{griffiths}.
For example, it does not provide a candidate for dark matter, which is estimated
to make up 25\% of the matter-energy density in the universe.
In addition, the running coupling constants of the electroweak and
strong forces do not intersect at a common energy scale. Another problem is that the force
of gravity does not appear in the \ac{SM}.

These points aside, the most important issue concerning
the Higgs sector of the \ac{SM} is the hierarchy problem.
The \ac{SM}
is accepted to be an effective field theory that describes
the elementary particles and their interactions well, up to an energy scale $\Lambda$ beyond which it breaks down.
This is the energy scale at which new physics must enter.
It is known that $\Lambda \leq \mathcal{O}(10^{19})$, the Planck scale, as quantum 
gravitational effects start to become important in that region. The energy scale $\Lambda$ enters
the corrections due to fermion and boson loops to the Higgs boson mass quadratically \cite{MSSM-carena-haber}:
\begin{equation}\label{eqn:mh_hierarchy}
\begin{split}
&M_{\PHiggs_{\text{SM}}}^2  = (M_H^2)_0 + \Delta M_{\PHiggs}^2,\\
&\Delta M_{\PHiggs}^2 \sim \mathcal{O}(\Lambda^2).
\end{split}
\end{equation}
Thus assuming that there is no new physics all the way up to
the Planck scale the observation $M_{\PHiggs}=125\,\GeV$ would only be compatible with $\Lambda$ of order $10^{19}\,\GeV$
if there was extreme fine-tuning of the bare Higgs boson mass $(M_{\PHiggs}^2)_0$.

@@

This can be used to define

@@

The remaining five degrees of freedom lead to five physical Higgs bosons: a charged

@@

and the neutral scalars \PHiggslight and \PHiggs. These arise from the mixing of the

@@

m_{\PHiggsps}^2\sin{\beta}^2 + m_{\PZ}^2\text{cos}^2\beta & -(m_{\PHiggsps}^2+m_{\PZ}^2)\sin{\beta}\cos{\beta}\\
-(m_{\PHiggsps}^2+m_{\PZ}^2)\sin{\beta}\cos{\beta} & m_{\PHiggsps}^2\text{cos}^2\beta+m_{\PZ}^2\text{sin}^2{\beta} \end{pmatrix},

@@

m_{\PHiggs,\PHiggslight}^2 = \frac{1}{2}[m_{\PHiggsps}^2+m_{\PZ}^2 \pm \sqrt{(m_{\PHiggsps}^2+m_{\PZ}^2)^2-4m_{\PZ}^2m_{\PHiggsps}^2\text{cos}^2 2\beta}].

@@

\text{cos}^2(\beta-\alpha) = \frac{m_{\PHiggslight}^2(m_{\PZ}^2-m_{\PHiggslight}^2)}{m_{\PHiggsps}^2(m_{\PHiggs}^2-m_{\PHiggslight}^2)}.

@@

The tree-level couplings of the neutral Higgs bosons to bosons and fermions 
are modified with respect to the couplings of the Higgs boson in the \ac{SM} by the multiplicative factors
given in table \ref{tab:mssm_couplings} \cite{YR4}.

@@

The dominant neutral MSSM Higgs boson production processes at hadron colliders
are slightly different from the \ac{SM} Higgs boson production processes. 
Because the pseudoscalar A does not couple to the
vector bosons, and in the decoupling limit the coupling of \PHiggs to vector
bosons is suppressed with respect to the \ac{SM} expectation, \ac{VBF} and \PW or \PZ associated
production ($\PW\PHiggs$ or $\PZ\PHiggs$) are not as important as for the production of the \ac{SM} Higgs boson. As in the \ac{SM}, gluon fusion production is 
the dominant production mode at low \tanb.
An example tree-level Feynman diagram for gluon fusion production, which proceeds via
a quark loop, is given in figure \ref{fig:production_mssm}a.
At low \tanb~values this loop is dominated by top quarks, while at 
high \tanb~the b-quark loop dominates. This is a result of the enhanced coupling to down-type fermions at high \tanb. 
Due to the negative top-bottom loop interference effect the gluon fusion cross section decreases with increasing \tanb~up
to around \tanb~$=3$, where the cross section starts to increase with increasing \tanb~again.
